---
title: "Vectors and lists"
output: 
  github_document:
    toc: true
editor_options:
  chunk_output_type: console
---

Working through [this section](https://jennybc.github.io/purrr-tutorial/bk00_vectors-and-lists.html) of the purrr tutorials

```{r}

library(tidyverse)

```
    
  
---------
  
  
Walking through section
  
## Atomic vectors
  
Atomic vectors are homogenous 
```{r}

v_log <- c(TRUE, FALSE, FALSE, TRUE)
v_int <- 1:4
v_doub <- 1:4 * 1.2
v_char <- letters[1:4]

```
  
### Excercises
  
1. Define the vectors above or similar. Use the family of is._() functions to confirm vector type, e.g. is.logical(). You will need to guess or look some of them up. Long-term, you may wish to explore the rlang::is_() family of functions.
```{r}

is.character(v_char)
is.numeric(v_doub)
is.integer(v_int)
is.logical(v_log)

```
  
2. What do is.numeric(), is.integer(), and is.double() return for the vectors that hold floating point number versus integers?
```{r}

is.numeric(v_doub)
is.numeric(v_int)

is.integer(v_doub) # FALSE
is.integer(v_int)

is.double(v_doub)
is.double(v_int) # FALSE

```

Can index with a logical vector or with integers  
```{r}

v_char[c(FALSE, FALSE, TRUE, TRUE)]
v_char[v_log]

```
  
```{r}

v_char[2:3]
v_char[-4]

```

### Exercises

1. What happens when you request the zero-th element of one of our vectors?
```{r}

v_char[0]

```
An atomic vector of length 0.
  
2. What happens when you ask for an element that is past the end of the vector, i.e. request x[k] when the length of x is less than k?
```{r}

v_doub[50]

```
  
3. We indexed a vector x with a vector of positive integers that is shorter than x. What happens if the indexing vector is longer than x?
```{r}

v_char[1:50]

```
  
4. We indexed x with a logical vector of the same length. What happen if the indexing vector is shorter than x?
```{r}

v_doub[c(TRUE, FALSE)]

```


----------  
  
## Coercion
  
Use as._() for explicit coerction
```{r}

as.integer(v_log)
as.numeric(v_int)
as.character(v_doub)

```
  
Implicit coercion can be triggered from other actions as well.
```{r}

v_doub_copy <- v_doub

v_doub_copy[3] <- "uhoh"
str(v_doub_copy)

```
  
Be intentional about type!
```{r}

big_plans <- rep(NA_integer_, 4)
str(big_plans)

big_plans[3] <- 5L # L implies double instead of int
str(big_plans)

```

### Exercises
  
1. Recall the hieararchy of the most common atomic vector types: logical < integer < numeric < character. Try to use the as.*() functions to go the wrong way. Call as.logical(), as.integer(), and as.numeric() on a character vector, such as letters. What happens?
```{r}

as.logical(letters[1:4])
as.integer(letters[1:4])
as.numeric(letters[1:4])

```
So can make anything a character vector, can make anything logical & integer a numeric vector, but (mostly) can't make a character vector logical, integer, or numeric. 
   
----------  
  
## Lists

Lists are useful if you want:

- An individual atom to be longer than length 1 (i.e. a sublist)
- Individual atoms of different type

```{r}

x <- list(1:3, c("four", "five"))
x

y <- list(letters[26:22], transcendental = c(pi, exp(1)), f = function(x) x^2)
y

```
  
### Exercises 
  
1. Make the lists x, y, and z as shown above. Use is._() functions to get to know these objects. Try to get some positive and negative results, i.e. establish a few things that x is and is NOT. Make sure to try is.list(), is.vector(), is.atomic(), and is.recursive(). Long-term, you may wish to explore the rlang::is._() family of functions.
```{r}

x
is.integer(x)
is.integer(x[[1]])
is.list(x)
is.vector(x)
is.atomic(x)
is.recursive(x)

```
  
  
----------
  
## List indexing
  
3 ways to index a list:

- With a single bracket
```{r}

x
x[1]
x[2]

```
This will return the specified element, but will always be a list

- With double square brackets
```{r}

x[[1]]
y[["transcendental"]]

```
This returns the "naked" element.

- With the $
```{r}

y$transcendental

```
Must specify the component by name with $.
  
### Exercises
  
1. Use [, [[, and $ to access the second component of the list z, which bears the name “transcendental”. Use the length() and the is.*() functions explored elsewhere to study the result. Which methods of indexing yield the same result vs. different?
```{r}

y
single <- y[2]
double <- y[[2]]
dollar <- y$transcendental

length(single)
length(double)
length(dollar)

is.list(single)
is.list(double)
is.atomic(dollar)

```


2. Put the same data into an atomic vector and a list:
```{r}

my_vec <- c(a = 1, b = 2, c = 3)
my_list <- list(a = 1, b = 2, c = 3)

```
Use [ and [[ to attempt to retrieve elements 2 and 3 from my_vec and my_list. What succeeds vs. fails? What if you try to retrieve element 2 alone? Does [[ even work on atomic vectors? Compare and contrast the results from the various combinations of indexing method and input object
```{r, error=TRUE}

my_vec[c(2, 3)]
my_vec[[c(2, 3)]]

```
  
  
----------
  
  
## Vectorized operations
  
Brute force way to square elements in a vector:
```{r}

n <- 5
res <- rep(NA_integer_, n)

for(i in seq_len(n)){
  res[i] <- i ^ 2
}

res

```
  
The R way:
```{r}

seq_len(n) ^ 2

```
  
But this is not possible with lists  
```{r, error=TRUE}

l_doub <- as.list(v_doub)

exp(l_doub)

```
  
So how do we apply a function to each element of a list? By using purrr:map()!
```{r}

map(l_doub, exp)

```





