---
title: "Intro to map() walkthrough"
subtitle: ""
output:
  github_document:
    toc: true
allow_html: true
editor_options:
  chunk_output_type: console
fig.width: 10
fig.height: 7
---


# Setup

```{r}

library(tidyverse)
library(repurrrsive)
library(magrittr)

theme_set(theme_light())

knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  fig.width = 10, 
  fig.height = 7
  )

```

----------

`purrr:map()` is used to apply a function to each element of a list

```{r}

map(1:5, sqrt)

```

So for each element 1, 2, ..., 5, the `sqrt()` function is applied and a list of each `sqrt(.)` is returned.

# Name and position shortcuts

`purrr:map` functions have shortcuts that allow the extraction of elemennts based on name or position.

```{r}

str(got_chars[1:4], list.len = 3)

```

For each sublist in `got_chars`, there is a `$name` element

```{r}

map(got_chars[1:4], "name")

```

An integer can be provided instead of a name
```{r}

map(got_chars[1:4], 3)

```

And it's easy to use this functionality in a pipe
```{r}

got_chars[1:4] %>% 
  map("name")

```

## Exercises

1. Use names() to inspect the names of the list elements associated with a single character. What is the index or position of the playedBy element? Use the character and position shortcuts to extract the playedBy elements for all characters.
```{r}

names(got_chars[[1]])

```

playedBy is the 18th element.

```{r}

got_chars %>% 
  map("playedBy")

```

```{r}

got_chars %>% 
  map(18)

```

2. What happens if you use the character shortcut with a string that does not appear in the lists’ names?

```{r}

got_chars %>% 
  map(";alskdjf")

```

A NULLED list of equal length to the outermost list is returned.

3. What happens if you use the position shortcut with a number greater than the length of the lists?

```{r}

got_chars %>% 
  map(50)

```

Same as with the non-name.

4. What if these shortcuts did not exist? Write a function that takes a list and a string as input and returns the list element that bears the name in the string. Apply this to got_chars via map(). Do you get the same result as with the shortcut? Reflect on code length and readability.

```{r}

map_knockoff_str <- function(l, s) l[[s]]
  
sapply(got_chars[1:4], map_knockoff_str, s = "name")

```

Less readable! Redundant if it's such a common use case.

5. Write another function that takes a list and an integer as input and returns the list element at that position. Apply this to got_chars via map(). How does this result and process compare with the shortcut?

```{r}

map_knockoff_int <- function(l, i) l[[i]]
  
sapply(got_chars[1:4], map_knockoff_int, i = 3)

```

----------

# Type-specific map

Variants of map() like map_chr() will return from a map call the exact type of object you want.
```{r}

got_chars %>% 
  map_chr("name")

```

A character vector is returned instead of a list

## Exercises

1. For each character, the second element is named “id”. This is the character’s id in the API Of Ice And Fire. Use a type-specific form of map() and an extraction shortcut to extract these ids into an integer vector.
```{r}

got_chars %>% 
  map_int("id")

```

2. Use your list inspection strategies to find the list element that is logical. There is one! Use a type-specific form of map() and an extraction shortcut to extract these values for all characters into a logical vector.
```{r}

got_chars %>% 
  map_lgl("alive")

```

Like how `map_chr()` returns a character vector, `map_lgl()` returns a logical vector.

----------

# Extract multiple values

Sometimes instead of just retrieving a single value from each list element (e.g. "name") we want multiple values (e.g. "name", "culture", and "gender").

For a single character, we can use 
```{r}

got_chars[[3]][c("name", "culture", "gender")]

```

And with `map()`, we have a few options for multiple value extraction:

- **Using `[`**

```{r}

got_chars[1:4] %>% 
  map(`[`, c("name", "culture", "gender"))

```

- **Using `extract()`**

```{r}

got_chars[1:4] %>% 
  map(extract, c("name", "culture", "gender"))

```

## Exercises

1. Use your list inspection skills to determine the position of the elements named “name”, “gender”, “culture”, “born”, and “died”. Map `[` or `magrittr::extract()` over users, requesting these elements by position instead of name.
```{r}

got_chars[[1]] %>% 
  names(.)

```

```{r}

got_chars[1:4] %>% 
  map(extract, c(3, 4, 5, 6, 7))

```

----------

# Data frame output

`map_dfr()` allows us to extract multiple values *and* stack the results in a data frame instead of keeping it as a nested list.
```{r}

got_chars[1:4] %>% 
  map_dfr(extract, c("name", "culture", "gender"))

```

The variable types in the resulting df are automatically type converted. This can be problematic sometimes, so it's good practice to explicitly type variables.
```{r}

got_chars %>% {
  
  tibble(
    name = map_chr(., "name"),
    culture = map_chr(., "culture"),
    gender = map_chr(., "gender"),
    id = map_int(., "id"),
    born = map_chr(., "born"),
    alive = map_lgl(., "alive")
  )
}

```

The curly brackets `{}` call prevents `got_chars` from being passed in as the first argument of `tibble()`.

## Exercises

1. Use `map_dfr()` to create the same data frame as above, but indexing with a vector of positive integers instead of names.
```{r}

got_chars[1:4] %>%
  map_dfr(extract, c(3, 5, 4, 2, 6, 8))

```

----------


