---
title: "Relationships to base and plyr"
subtitle: ""
output:
  github_document:
    toc: true
allow_html: true
editor_options:
  chunk_output_type: console
fig.width: 10
fig.height: 7
---

```{r}

library(tidyverse)
library(repurrrsive)
library(gapminder)

theme_set(theme_light())

knitr::opts_chunk$set(
  cache = FALSE,
  message = FALSE, 
  warning = FALSE, 
  fig.width = 10, 
  fig.height = 7
  )

```

# lapply() vs purrr:map()

These are list in, list out functions.

- **Base**
```{r}

got_chars[1:3] %>% 
  lapply(function(x) x[["name"]])

```

- **purrr**
```{r}

a <- got_chars[1:3] %>% 
  map("name")

```

So purrr knows that "name" is a named element in the list got_chars[1:3]. It returns a list length 3 with each character's name.

----------

# vapply() vs map_*()

vapply() requires a template for the return value, while purrr has map_*() functions like map_chr() 

- **Base**
```{r}

got_chars[1:3] %>% 
  vapply(function(x) x[["name"]],
         character(1))

```

- **purrr*
```{r}

got_chars[1:3] %>% 
  map_chr("name")

```

So again, purrr knows that "name" is a named element in the list got_chars[1:3]; however, this time it returns a character vector with the names because map_chr() is used instead of just map().

----------

# map_dfr()

This function is list in, data frame out

- **Base**
```{r}

l <- got_chars[23:25] %>% 
  lapply("[", c("name", "playedBy"))

mat <- do.call(rbind, l)

(df <- as.data.frame(mat, stringsAsFactors = FALSE))

```

- **purrr**
```{r}

got_chars[23:25] %>% 
  map_dfr(`[`, c("name", "playedBy"))

```

----------

# mapply() vs map2(), pmap()

Functions used for mapping over 2+ vectors/lists.

- **Base**
```{r}

nms <- vapply(
  got_chars[16:18],
  `[[`,
  character(1),
  "name"
  )

birth <- vapply(
  got_chars[16:18],
  `[[`,
  character(1),
  "born"
  )

mapply(
  function(x, y) paste(x, "was born", y),
  nms, birth
  )

```

- **purrr**
```{r}

nms <- got_chars[16:18] %>% 
  map_chr("name")

birth <- got_chars[16:18] %>% 
  map_chr("born")

map2_chr(nms, birth, ~ paste(.x, "was born", .y))

```

----------

# aggregate() vs dplyr::summarise()

```{r}

(mini_gap <- gapminder %>% 
  filter(country %in% c("Canada", "Germany"), year > 2000) %>% 
  droplevels())

```

Single variable summary:

- **Base**
```{r}

aggregate(lifeExp ~ country, mini_gap, mean)

```

- **tidyverse**
```{r}

mini_gap %>% 
  group_by(country) %>% 
  summarise(lifeExp = mean(lifeExp))

```

Multiple variable summary:

- **Base**
```{r}

aggregate(cbind(lifeExp, gdpPercap) ~ country, mini_gap, mean)

```

- **tidyverse**
```{r}

mini_gap %>% 
  group_by(country) %>% 
  summarise_at(vars(lifeExp, gdpPercap), mean)

```

Bivariate summary of two variables for each country:

- **Base**
```{r}

by_obj <- by(gapminder, gapminder$country, function(df) cor(df$lifeExp, df$year))

head(by_obj)

```

- **tidyverse**
```{r}

gapminder %>% 
  group_by(country) %>% 
  summarise(cor = cor(lifeExp, year))

```

----------

# by() vs tidyr::nest()

Goal: fit a linear model of life expectancy vs year

- **Base**
```{r}

by_obj <- by(gapminder,
             gapminder$country,
             function(df) lm(lifeExp ~ year, data = df)
             )

str(by_obj[1:2], max.level = 1)

by_obj$Afghanistan

```

Country information lives inside by_obj, a nested list, and finding information about the country like its continent is now quite difficult.

```{r}

o_countries <- as.character(unique(gapminder$country[gapminder$continent == "Oceania"]))

by_obj[names(by_obj) %in% o_countries]

```

In the tidyverse, we create a nested data frame where each row is a country and the last column contains the linear model data for that country

- **tidyverse**
```{r}

nested_df <- gapminder %>% 
  group_by(country, continent) %>% 
  nest() %>% 
  mutate(fit = map(data, ~ lm(lifeExp ~ year, data = .x)))

str(nested_df$fit[1:2], max.level = 1)

```

The values in the data column are nested data frames
```{r}

nested_df$data[1]

```

And the values in fit are linear model objects
```{r}

nested_df$fit[1]

```

Having the data and fit objects makes a few things much easier:

- **Filtering**
```{r}

nested_df %>% 
  filter(continent == "Oceania") %>% 
  .$fit

```

- **Extracting and including fit values**
```{r}

nested_df %>% 
  mutate(
    coefs = map(fit, coef),
    intercept = map_dbl(coefs, 1),
    slope = map_dbl(coefs, 2)
    ) %>% 
  select(country, continent, intercept, slope)

```

coef() is a function that grabs coefficients from a model object. So the columns are created with maps that grab the coefficients from the fit column.

Below it's broken down into a mapping of the fits to get coefficients, then a mapping of that returned list to get the numeric values.

- **Get a list of coefficients**
```{r}

coef(nested_df$fit[[1]])

map(nested_df$fit, coef) %>% head(3)

```

- **Get each intercept**
```{r}

map(nested_df$fit, coef) %>% head(3) %>% map_dbl(., 1)

```

- **Get each slope**
```{r}

map(nested_df$fit, coef) %>% head(3) %>% map_dbl(., 2)

```



----------








