---
title: "Explore the example lists: Wes Anderson, GoT, GitHub"
subtitle: ""
output:
  github_document:
    toc: true
allow_html: true
editor_options:
  chunk_output_type: console
fig.width: 10
fig.height: 7
---

# Setup

```{r}

library(tidyverse)
library(repurrrsive)

theme_set(theme_light())

knitr::opts_chunk$set(
  cache = TRUE,
  message = FALSE, 
  warning = FALSE, 
  fig.width = 10, 
  fig.height = 7
  )

```

----------

# Inspect and explore

It's important to understand the structure of a list when we're mapping through it, especially

- The length
- Whether components are homogenous
- Length/names/types of each nested list

## Indexing

There are 3 ways to index a list

```{r}

x <- list(a = "a", b = 2)

```


1. With `$`
```{r}

x$a

```

2. With `[[`
```{r}

x[["a"]]

```

This returns the naked element

3. With [
```{r}

x["a"]

```

This returns a list

## str()

`str()` helps with basic inspection. Especially useful arguments are:

- `list.len`: Max # of elements to display within a level
- `max.level`: Max level of nesting applied in display

----------

# Wes Anderson color palettes

```{r}

str(wesanderson)

```

A list of 15 different pallettes, each named and each containing 4 character values.

----------

# GoT POV characters

```{r}

str(got_chars, list.len = 3) # list.len = 3 limits # elements shown from first level sublist to 3

```

A list of 30 characters, each character with 18 named sublists containing information on them.

----------

# GitHub users and repos

```{r}

str(gh_users, max.level = 1) 

```

----------

# Exercises


1. Read the documentation on `str()`. What does `max.level` control? Apply `str()` to `wesanderson` and/or `got_chars` and experiment with `max.level = 0`, `max.level = 1`, and `max.level = 2`. Which will you use in practice with deeply nested lists?
```{r}

str(got_chars, max.level = 0)

```

```{r}

str(got_chars, max.level = 1)

```

```{r}

str(got_chars, max.level = 2)

```

2. What does the `list.len` argument of `str()` control? What is its default value? Call `str()` on got_chars and then on a single component of `got_chars` with `list.len` set to a value much smaller than the default. What range of values do you think you’ll use in real life?
```{r}

str(got_chars, list.len = 1)

```

```{r}

str(got_chars, list.len = 5)

```


3. Call `str()` on got_chars, specifying both max.level and list.len.
```{r}

str(got_chars, max.level = 2, list.len = 5)

```

4. Call `str()` on the first element of `got_chars`, i.e. the first Game of Thrones character. Use what you’ve learned to pick an appropriate combination of `max.level` and `list.len`.
```{r}

theon <- got_chars[[1]]

str(theon, max.level = 1, list.len = 5)

```



